Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-07 ms.tex
zetaCrit figure

\begin{abstract}
  We calculate steady-state, one-dimensional hydrodynamic profiles of
  hot gas in slowly accreting (``quiescent") galactic nuclei for a range of
  central black hole masses $M_{\bullet}$, parameterized gas heating
  rates, and observationally-motivated stellar density profiles.  Mass
  is supplied to the circumnuclear medium by stellar winds, while gas
  is heated primarily by stellar winds, supernovae, and black hole
  feedback.  Analytic estimates are derived for the stagnation radius
  (where the radial velocity of the gas passes through zero) and the
  black hole accretion rate, $\dot{M}$, as a function of $M_{\bullet}$ and the gas heating efficiency, the latter being related to the star-formation history.  We assess the conditions under which radiative instabilities develop, separately in the case
  of a single burst of star formation and for the average star formation history predicted by cosmological simulations for galaxies hosting black holes of a given mass.  By combining a sample of measured nuclear X-ray luminosities, $L_{X}$, of nearby quiescent galactic nuclei with our results for $\dot{M}(M_{\bullet})$ we show that most nuclei are consistent with accreting in a steady-state, thermally-stable manner.  However, thermally-stable accretion cannot explain the short average growth times of low mass black holes in the local universe, which must instead result from gas being fed in from
  large radii, due either to galaxy mergers or thermal instabilities acting on larger, galactic scales.  Our results have implications for attempts to constrain the occupation fraction of SMBHs in low mass galaxies using the mean
  $L_{\rm X}-M_{\bullet}$ correlation, as well as the predicted diversity of the
  circumnuclear densities encountered by relativistic outflows from
  tidal disruption events.
\end{abstract}

@@

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & black hole mass \\
$\tilde{v}_{w}$ & total heating parameter, including minimum heating
rate from stellar and black hole velocity dispersion \\
$v_{w}$ & total heating parameter, excluding minimum heating rate from
velocity dispersion \\
$\sigma_0$ & stellar velocity dispersion (assumed to be radially constant)\\
$\zeta$ & heating parameter, $\zeta \equiv \sqrt{1 + (v_w/\sigma_0)^2}$ (eq.~[\ref{eq:zetacrit}])\\
$r_{s}$ & stagnation radius, where gas radial velocity goes to zero \\
$r_{\rm inf}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\ 
$r_{\rm Ia}$ & radius interior to which SN Ia are infrequent compared to the dynamical timescale (eq.~[\ref{eq:rIa}]) \\ 
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds, $q \propto \rho_{\star}$ (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar surface brightness
profile interior to the break radius \\
$\delta$ & power-law slope of the 3D stellar density profile inside of
the break radius, $\delta \equiv \Gamma+1$.\\
$\dot{M}$ & mass accretion rate  \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & `equilibrium' accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$t_{\rm h}$ & Hubble timescale \\
$\nu$ & gas density power-law slope at the stagnation radius
(eq.~[\ref{eq:densSlope}]) \\
\hline
\label{table:definitions}  
\end{tabular}
%\begin{tablenotes}
%\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

We fix $\beta = 2$ because our results are
generally not sensitive to the properties of the gas flow on radial
scales $\gtrsim r_b$. The presence of the break radius $r_{b}$ is,
however, necessary to obtain a converged steady state for some regions
of our parameter space.  We consider solutions calculated for up to
four values of $\rb$: 50 pc, 100 pc, 200 pc, and 400 pc, motivated by
the range of break radii from the \citet{LauerFaber+:2007a}
sample.\footnote{For the core galaxies the break radius follows the
  scaling relationship $\rb\sim 106 \, (\Mbheight)^{0.39}$ pc, with
  scatter of approximately one dex.  Most the cusp galaxies have $\rb$
  between 100 and 1000 pc, without a clear trend with $\Mbh$.}  In
practice we calculate solutions for a range of $\rb$ only in cases
when its precise value is expected to have a significant affect on the
solution, in particular for high SMBH masses and low heating rates,
for which the stagnation radius approaches large radii. We also
neglect value of $r_b$ which would give unphysically large bulges for
a given $\Mbh$.

@@

Radiative cooling contributes an additional term to equation (\ref{eq:dsdt}) of the form
\begin{equation}
\dot{q}_{\rm rad} = -\Lambda(T)n^{2},
\label{eq:qdot_rad}
\end{equation}
where $n = \rho/\mu m_p$ is the gas number density and $\Lambda(T)$ is
the cooling function.  We neglect radiative cooling in our baseline
calculations, despite the fact that this is not justified when the
wind heating $\vwO$ is low or if the mass return rate $\eta$ is high.
However, once radiative cooling becomes important compared to other
sources of heating and cooling, its presence leads to thermal
instability (e.g.~\citealt{McCourt+12}, \citealt{Li&Bryan14a}) that cannot be accounted for
by our 1D time-independent model.  A main goal of this work is to use
solutions neglecting radiation to ascertain over what range of
physical conditions cooling instabilities are most likely to develop
($\S\ref{sec:instability}$).

@@

In the opposite limit of weak heating ($v_w\lsim \sigma_0$) the
stagnation radius moves to large radii, approaching the break radius
$\rb$ in the stellar density profile, formally implying that almost
entire ISM of the galaxy is accreting.  In particular, in Appendix
\ref{app:be} we show that $\rs \Rightarrow \rb$ for heating below a
critical threshold \be \zeta\equiv \sqrt{1 + \frac{v_w}{\sigma_0}^2} <
\zeta_c \approx \left(\frac{\rb}{\rsoi}\right)^{0.5(1-\Gamma)}.
\label{eq:zetacrit}
\ee This result makes intuitive sense: gas is supplied to the nuclear
by stars which are gravitationally bound to the black hole, so
outflows are possible only if the specific heating rate $\sim
v_{w}^{2}$ significantly exceeds the specific gravitational binding
energy.  Condition (\ref{eq:zetacrit}) approximately corresponds to
the condition that the wind heating parameter be less than the local escape
speed from near the break radius, $v_{\rm esc} \sim (GM_{\star}|_{r_{\rm
    b}}/r_b)^{1/2}$, where $M_{\star} \propto r^{2-\Gamma}$.

@@

Radiative cooling usually has its greatest impact near or external to
the stagnation radius, where the gas resides in near hydrostatic
balance.  If radiative cooling becomes important, it can qualitatively
alter key features of the accretion flow. Initially hydrostatic gas
that cools at constant pressure $P \propto nT$ at the rate
$\dot{q}_{\rm rad} = n^{2}\Lambda \propto T^{-2.7}$
(eq.~[\ref{eq:qdot_rad}]; see below) is thermally unstable
(e.g., \citealt{McCourt+12}, \citealt{Li&Bryan14a}), potentially resulting
in the formation of a multi-phase medium and enhanced star formation (\citealt{Li&Bryan14b}).
Even if the hot plasma of the CNM does not condense into cold clouds,
the loss of pressure can temporarily increase the black hole accretion
rate.  When coupled to feed-back processes which result from such enhanced accretion, this can lead to
time-dependent limit cycle behavior
(e.g.~\citealt{Ciotti&Ostriker07}; \citealt{Ciotti+10};
\citealt{Yuan&Li11}, \citealt{Gan+14}), which is also inconsistent
with our assumption of a steady, single-phase flow.

@@

To within a constant of order unity, equation (\ref{eq:cooling2}) also
equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n
kT/2 \mu)/\dot{q}_{\rm rad}$ to the free-fall time $t_{\rm ff}$ at the
stagnation radius, as can be derived using the equality
\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \lesssim 10 t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well (see also \citealt{Li&Bryan14a}, \citealt{Li&Bryan14b}).

@@

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm inf}$).  However, for low heating the
stagnation radius diverges above the analytic estimate, approaching
the stellar break radius $r_b \gg r_{\rm inf}$.  This divergence
occurs approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which coincide with the Bernoulli parameter of the gas decreasing below unity near the stagnation radius (Appendix \ref{app:be}).  This divergence of the stagnation radius to large radii occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.8$) with large break radii, explaining the behavior of the two core galaxies shown in Fig.~\ref{fig:stag} as vertically aligned orange triangles.  These solutions correspond to the same black hole mass
($\Mbh = 10^8 \Msun$) and heating rate ($v_w = 600$ km s$^{-1}$) but assume different break radii of $r_{b} =$ 25, 50 pc (from top to
bottom).

@@

Fig.~\ref{fig:cooling} shows the ratio of wind heating to radiative
cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$
(eq.~[\ref{eq:cooling2}] and surrounding discussion) as a function of
radius for our fiducial solutions shown in Figure \ref{fig:profiles}.
For high heating rates of $v_{w} = 600$ and $1200$ km s$^{-1}$ cooling
is unimportant across all radii, while for $v_{w} = 300$ km s$^{-1}$
we see that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ and $\tcool/\tff$
can be less than unity, depending on the wind mass loss parameter
$\eta$.  Because at the stagnation radius $\dot{q}_{\rm
  heat}/|\dot{q}_{\rm rad}|$ reaches within a factor
of two of its minimum across the entire grid, the value of $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ can be taken as diagnostic of thermal instability at all radii.

@@

In addition to our solutions which neglect thermal conduction and radiative cooling, we explore a smaller set of simulations including these effects explicitly, as marked by dagger symbols in Table \ref{table:models}.  For solutions which are far from being thermally unstable when cooling is neglected (e.g., $M_{\bullet} = 10^{8}M_{\odot}$, $v_{w} = 600$ km s$^{-1}$, $r_{b} = 100$ pc, $\eta = 0.2$, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| = 76$), the inclusion of radiative cooling has little effect on the key properties of the flow, e.g., stagnation radius, mass accretion rate, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$.  However, for solutions which are already marginally thermally stable, including radiative losses acts to decrease the value of $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ by a greater fractional amount.  We calculated solutions with radiative cooling for all cusp galaxies with $v_w=300$ km s$^{-1}$ and for all core galaxies with $v_w=600 $ km s$^{-1}$, each for three different values of $\eta = 0.02, 0.2, 0.6$, spanning the physical range of expected mass loss rates.  Solutions found to be unstable for all values of $\eta$ are marked in boldface as ``TI" in Table \ref{table:models}, with the unstable values of $\eta$ provided in the final column.  In contrast, changes in the solution when thermal conduction was included were always found to be minor for our choice of the saturation parameter $\phi = 0.1$.  Conduction does, however, offset the effects of radiative cooling, making the flow more thermally stable (e.g.~\citealt{Zakamska&Narayan03}).

\section{Heating Sources}
\label{sec:heating}

@@

However, even if SN Ia are rare near the stagnation radius, SN Ia may cap the
SMBH accretion rate by periodically blowing gas out of the nucleus of
low mass galaxies.  Between successive SNe, stars release a
gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$
interior to the Ia radius, which is locally gravitationally bound to the SMBH
by an energy $E_{\rm bind} \sim M_{\rm g}\sigma_0^{2}\mathcal{F}^{2}$, where typically $\mathcal{F} 
\lesssim$ few.  From the above definitions we find that 
\be
\frac{E_{\rm Ia}}{E_{\rm bind}} \sim \frac{(v_{w}^{\rm Ia})^{2}}{2\mathcal{F}^{2}\sigma_0^{2}}.  
\label{eq:blowout}
\ee
Hence, for low mass BHs with $\sigma_0 \ll v^{\rm Ia}_{w}$, SN Ia are capable of dynamically clearing out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Even when heating is sufficiently weak that the stagnation radius formally exceeds $r_{\rm Ia}$, the SMBH accretion rate is still potentially limited to a value
\begin{eqnarray}
\frac{\dot{M}_{\rm Ia}}{\dot{M}_{\rm edd}} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{r_{\rm Ia}}{r_{\rm inf}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{-0.2}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{-0.6}   & \text{cusp},
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
obtained substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}])
for $r_{\rm s}$ in the derivation leading to our estimate of $\dot{M}$
(eq.~[$\ref{eq:eddr_analytic}$]).  
%Note, however, that for low mass
%SMBHs, $\dot{M}_{\rm Ia}$ generally exceeds the maximum accretion rate
%for a thermally stable flow $\dot{M}_{\rm TI}$
%(eq.~[\ref{eq:Mdotmax}]), implying that dynamical `blow-out' from SNe
%Ia cannot alone prevent cooling instabilities in the nuclei of low
%mass galaxies.

The deep gravitational potential wells of high mass galaxies $\sigma_0
\gg  v^{\rm Ia}_{w}$ prevent SN Ia from dynamically clearing out gas
in these systems.  Equation (\ref{eq:eddr_Ia}) nevertheless represents
a cap on the accretion rate in practice because if other source of
heating are ineffective, then the Ia heating rate (eq.~[\ref{eq:vIa}])
is usually sufficiently high to prevent the stagnation radius
(calculated including the heating from SN Ia) from substantially
exceeding $r_{\rm Ia}$. Stated another way, if the stagnation radius moves inwards from $\rIa$, then decreased heating will force it outwards again. However, if the stagnation radius moves well outside of $\rIa$, then the high level of Ia heating will force it inwards.

>>>>

\begin{abstract}
  We calculate steady-state, one-dimensional hydrodynamic profiles of
  hot gas in ``quiescent" galactic nuclei for a range of
  central black hole masses $M_{\bullet}$, parameterized gas heating
  rates, and observationally-motivated stellar density profiles.  Mass
  is supplied to the circumnuclear medium by stellar winds, while gas
  is heated primarily by stellar winds, supernovae, and black hole
  feedback.  Analytic estimates are derived for the stagnation radius
  (where the radial velocity of the gas passes through zero) and the
  black hole accretion rate, $\dot{M}$, as a function of $M_{\bullet}$ and the gas heating efficiency, the latter being related to the star-formation history.  We assess the conditions under which radiative instabilities develop, separately in the case
  of a single burst of star formation and for the average star formation history predicted by cosmological simulations for galaxies hosting black holes of a given mass.  By combining a sample of measured nuclear X-ray luminosities, $L_{X}$, of nearby quiescent galactic nuclei with our results for $\dot{M}(M_{\bullet})$ we show that most nuclei are consistent with accreting in a steady-state, thermally-stable manner.  However, thermally-stable accretion cannot explain the short average growth times of low mass black holes in the local universe, which must instead result from gas being fed in from
  large radii, due either to galaxy mergers or thermal instabilities acting on larger, galactic scales.  Our results have implications for attempts to constrain the occupation fraction of SMBHs in low mass galaxies using the mean
  $L_{\rm X}-M_{\bullet}$ correlation, as well as the predicted diversity of the
  circumnuclear densities encountered by relativistic outflows from
  tidal disruption events.
\end{abstract}

@@

\begin{table*}
\begin{threeparttable}
%\centering
\begin{minipage}{18cm}
\caption{Definitions of commonly used variables}
\begin{tabular}{ll}
\hline
{Variable} & {Definition} \\
\hline
$M_{\bullet}$ & black hole mass \\
$\tilde{v}_{w}$ & total heating parameter, including minimum heating
rate from stellar and black hole velocity dispersion \\
$v_{w}$ & total heating parameter, excluding heating rate from
velocity dispersion \\
$\sigma_0$ & stellar velocity dispersion--assumed to be radially constant.\\
$\zeta$ & heating parameter, $\zeta \equiv \sqrt{1 + (v_w/\sigma_0)^2}$ (eq.~[\ref{eq:zetacrit}])\\
$r_{s}$ & stagnation radius (where gas radial velocity goes to zero) \\
$r_{\rm inf}$ & radius of sphere of influence (eq.~[\ref{eq:rsoi}]) \\
$r_{\rm b}$ & outer break radius of stellar density profile \\ 
$r_{\rm Ia}$ & radius interior to which Ia SN are infrequent compared to the dynamical/inflow timescale (eq.~[\ref{eq:rIa}]) \\ 
$\rho_{\star}(r)$ & 3D radial stellar density profile \\
$\rho(r)$ & gas density of CNM \\
$M_{\star}(r)$ & total enclosed stellar mass inside radius $r$ \\
$M_{\rm enc}(r)$ & total enclosed mass inside radius $r$ (SMBH + stars) \\
$q(r)$ & mass source term due to stellar winds (eq.~[\ref{eq:q}]) \\
$\eta$ & parameter setting normalization of mass input from stellar winds (eq.~[\ref{eq:q}]) \\
$\tau_{\star}$ & age of stellar population, in case of a single burst of star formation \\
$\Gamma$ & power-law slope of radial stellar surface brightness
profile interior to the break radius \\
$\delta$ & power-law slope of the 3d stellar density profile inside of
the break radius. $\delta=\Gamma+1$.\\
$\dot{M}$ & mass accretion rate  \\
$\dot{M}_{\rm Ia}$ & maximum mass accretion rate as limited by SN Ia (eq.~[\ref{eq:eddr_Ia}]) \\
$\dot{M}_{\rm C}$ & `equilibrium' accretion rate set by Compton heating (eq.~[\ref{eq:MdotC}]) \\
$\dot{M}_{\rm TI}$ & maximum accretion rate for thermally stable accretion (eq.~[\ref{eq:Mdotmax}]) \\
$\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ & ratio of external heating (stellar winds, SN Ia, MSPs) to radiative cooling (eq.~[\ref{eq:cooling2}]) \\
$t_{\rm h}$ & Hubble timescale \\
$\nu$ & gas density power-law slope at the stagnation radius
(eq.~[\ref{eq:densSlope}]) \\
\hline
\label{table:definitions}  
\end{tabular}
%\begin{tablenotes}
%\end{tablenotes}
\end{minipage}
\end{threeparttable}

@@

We fix $\beta = 2$ (a typical value ${\bf AG beta=1.5 may be more
  typical...but it still may not matter...}$) because our results are
generally not sensitive to the properties of the gas flow on radial
scales $\gtrsim r_b$. The presence of the break radius $r_{b}$ is,
however, necessary to obtain a converged steady state for some regions
of our parameter space.  We consider solutions calculated for up to
four values of $\rb$: 50 pc, 100 pc, 200 pc, and 400 pc, motivated by
the range of break radii from the \citet{LauerFaber+:2007a}
sample.\footnote{For the core galaxies the break radius follows the
  scaling relationship $\rb\sim 106 \, (\Mbheight)^{0.39}$ pc, with
  scatter of approximately one dex.  Most the cusp galaxies have $\rb$
  between 100 and 1000 pc, without a clear trend with $\Mbh$.}  In
practice we calculate solutions for a range of $\rb$ only in cases
when its precise value is expected to have a significant affect on the
solution, in particular for high SMBH masses and low heating rates,
for which the stagnation radius approaches large radii. We also
neglect value of $r_b$ which would give unphysically large bulges for
a given $\Mbh$.

@@

Radiative cooling contributes an additional term to equation (\ref{eq:dsdt}) of the form
\begin{equation}
\dot{q}_{\rm rad} = -\Lambda(T)n^{2},
\label{eq:qdot_rad}
\end{equation}
where $n = \rho/\mu m_p$ is the gas number density and $\Lambda(T)$ is
the cooling function.  We neglect radiative cooling in our baseline
calculations, despite the fact that this is not justified when the
wind heating $\vwO$ is low or if the mass return rate $\eta$ is high.
However, once radiative cooling becomes important compared to other
sources of heating and cooling, its presence leads to thermal
instability (e.g.~\citealt{McCourt+12}) that cannot be accounted for
by our 1D time-independent model.  A main goal of this work is to use
solutions neglecting radiation to ascertain over what range of
physical conditions cooling instabilities are most likely to develop
($\S\ref{sec:instability}$).

@@

In the opposite limit of weak heating ($v_w\lsim \sigma_0$) the
stagnation radius moves to large radii, approaching the break radius
$\rb$ in the stellar density profile, formally implying that almost
entire ISM of the galaxy is accreting.  In particular, in Appendix
\ref{app:be} we show that $\rs \Rightarrow \rb$ for heating below a
critical threshold \be \zeta\equiv \sqrt{1 + \frac{v_w}{\sigma_0}^2} <
\zeta_c \approx \left(\frac{\rb}{\rsoi}\right)^{0.5(1-\Gamma)}.
\label{eq:zetacrit}
\ee This result makes intuitive sense: gas is supplied to the nuclear
by stars which are gravitationally bound to the black hole, so
outflows are possible only if the specific heating rate $\sim
v_{w}^{2}$ significantly exceeds the specific gravitational binding
energy.  Condition (\ref{eq:zetacrit}) approximately corresponds to
the condition that the wind heating parameter be less than the escape
speed from the break radius, $v_{\rm esc} \sim (GM_{\star}|_{r_{\rm
    b}}/r_b)^{1/2}$, where $M_{\star} \propto r^{2-\Gamma}$.

@@

Radiative cooling usually has its greatest impact near or external to
the stagnation radius, where the gas resides in near hydrostatic
balance.  If radiative cooling becomes important, it can qualitatively
alter key features of the accretion flow. Initially hydrostatic gas
that cools at constant pressure $P \propto nT$ at the rate
$\dot{q}_{\rm rad} = n^{2}\Lambda \propto T^{-2.7}$
(eq.~[\ref{eq:qdot_rad}]; see below) is thermally unstable
(\citealt{McCourt+12} and references therein), potentially resulting
in the formation of a multi-phase medium and enhanced star formation.
Even if the hot plasma of the CNM does not condense into cold clouds,
the loss of pressure can temporarily increase the black hole accretion
rate.  When coupled to feed-back processes which result from such enhanced accretion, this can lead to
time-dependent limit cycle behavior
(e.g.~\citealt{Ciotti&Ostriker07}; \citealt{Ciotti+10};
\citealt{Yuan&Li11}, \citealt{Gan+14}), which is also inconsistent
with our assumption of a steady, single-phase flow.

@@

To within a constant of order unity, equation (\ref{eq:cooling2}) also
equals the ratio of the gas cooling timescale $t_{\rm cool} \equiv (3n
kT/2 \mu)/\dot{q}_{\rm rad}$ to the free-fall time $t_{\rm ff}$ at the
stagnation radius, as can be derived using the equality

\begin{align}
\rho|_{r_{\rm s}}=\frac{3 q\tff|_{r_{\rm s}}}{2-\Gamma}.
\label{eq:rhors2}
\end{align}
that results by combining equations
\eqref{eq:mdot_analytic},\eqref{eq:mdot_gas}, and
\eqref{eq:rhostarrs}.  \citet{McCourt+12} argue cooling instability
develops in a hydrostatic atmosphere if $t_{\rm cool} \ll t_{\rm ff}$,
so equation (\ref{eq:cooling2}) represents a good proxy for
instability in this case as well.

@@

Our analytic estimates accurately reproduce the numerical results in
the high heating limit $\zeta \gg 1$ ($v_w \gg \sigma_0$; $r_{\rm s}
\lesssim r_{\rm inf}$).  However, for sufficiently low heating the
stagnation radius diverges above the analytic estimate, approaching
the stellar break radius $r_b \gg r_{\rm inf}$.  This divergence
occurs approximately when $\zeta < \zeta_c \propto (r_b/r_{\rm
  soi})^{0.5(\Gamma-1)}$ (eq.~[\ref{eq:zetacrit}]), which in Appendix
\ref{app:be} is shown to coincide with the Bernoulli parameter of the gas decreasing below unity near the stagnation radius.  This divergence of the stagnation radius to large radii hence occurs at a higher value of $\zeta$ in core galaxies ($\Gamma = 0.8$) with large break radii.  This dependence explains the behavior of the two core galaxies, shown as
orange triangles vertically aligned at the same value of
$\zeta$; the solutions correspond to the same black hole mass
($\Mbh = 10^8 \Msun$) and heating rate ($v_w = 600$ km s$^{-1}$) but are calculated assuming
different values of the break radius: $r_{b} =$ 25, 50 pc (from top to
bottom).

@@

Fig.~\ref{fig:cooling} shows the ratio of wind heating to radiative
cooling, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$
(eq.~[\ref{eq:cooling2}] and surrounding discussion) as a function of
radius for our fiducial solutions shown in Figure \ref{fig:profiles}.
For high heating rates of $v_{w} = 600$ and $1200$ km s$^{-1}$ cooling
is unimportant across all radii, while for $v_{w} = 300$ km s$^{-1}$
we see that $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ and $\tcool/\tff$
can be less than unity, depending on the wind mass loss parameter
$\eta$.  At the stagnation radius, the ratio $\dot{q}_{\rm
  heat}/|\dot{q}_{\rm rad}|$ reaches a value that lies within a factor
of two of its minimum across the entire grid; thus, the value of
$(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}}$ provides a
diagnostic of thermal instability for the entire solution.

@@

In addition to our standard set of solutions which neglect thermal conduction and radiative cooling, we explore a few test simulations including these effects, as marked by dagger symbols in Table \ref{table:models}.  For solutions which are far from being thermally unstable when cooling is neglected (e.g., $M_{\bullet} = 10^{8}M_{\odot}$, $v_{w} = 600$ km s$^{-1}$, $r_{b} = 100$ pc, $\eta = 0.2$, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| = 76$), including radiative cooling has little effect on the key properties of the flow (e.g., stagnation radius, mass accretion rate, $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$).  However, for solutions which are marginally stable (e.g., $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}| \sim $few), the inclusion of radiative acts to decrease the value of $\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|$ by a greater amount.  We ran solutions with radiative cooling for all cusp galaxies with $v_w=300$ km s$^{-1}$ and all core galaxies with $v_w=600 $ km s$^{-1}$, each for three different values of the mass input parameter $\eta = 0.02, 0.2, 0.6$, spanning the physical range.  Solutions found to be unstable for all values of $\eta$ are marked in boldface as ``TI" in Table \ref{table:models}.
By contrast, changes due to the inclusion of thermal conduction are always minor for our choice of the saturation parameter $\phi = 0.1$, although conduction is seen to offset the effects of radiative cooling, making the flow more thermally stable (e.g.~\citealt{Zakamska&Narayan03}).

\section{Sources of Heating}
\label{sec:heating}

@@

However, even if SN Ia are rare near the stagnation radius, SN Ia may cap the
SMBH accretion rate by periodically blowing gas out of the nucleus of
low mass galaxies.  Between successive SNe, stars release a
gaseous mass $M_{\rm g} \approx \eta M_{\star}\tau_{\rm Ia}/t_{\rm h}$
interior to the Ia radius, which is locally gravitationally bound to the SMBH
by an energy $E_{\rm bind} \sim M_{\rm g}\sigma_0^{2}\mathcal{F}^{2}$, where typically $\mathcal{F} 
\lesssim$ few.  From the above definitions we find that 
\be
\frac{E_{\rm Ia}}{E_{\rm bind}} \sim \frac{(v_{w}^{\rm Ia})^{2}}{2\mathcal{F}^{2}\sigma_0^{2}}.  
\label{eq:blowout}
\ee
Hence, for low mass BHs with $\sigma_0 \ll \mathcal{F}v^{\rm Ia}_{w}/2$, SN Ia are capable of dynamically clearing out gas from radii $\sim r_{\rm Ia} \gtrsim r_{\rm s}$.  Even when heating is sufficiently weak that the stagnation radius formally exceeds $r_{\rm Ia}$, the SMBH accretion rate is still potentially limited to a value
\begin{eqnarray}
\frac{\dot{M}_{\rm Ia}}{\dot{M}_{\rm edd}} &\approx& \frac{\eta M_{\bullet}}{\dot{M}_{\rm edd} t_{\rm h}}\left(\frac{r_{\rm Ia}}{r_{\rm inf}}\right)^{2-\Gamma} \approx \nonumber \\
 && \begin{cases}
    4.5 \times 10^{-4} M_{\bullet,8}^{-1.33}(\tau_{\star}/t_{\rm h})^{-0.2}
   & \text{core} \\
    2.2 \times 10^{-4} \Mbheight^{-0.84}(\tau_{\star}/t_{\rm h})^{-0.6}   & \text{cusp},
  \end{cases}
  \label{eq:eddr_Ia}
\end{eqnarray}
obtained substituting the Ia radius $r_{\rm Ia}$ (eq.~[\ref{eq:rIa}])
for $r_{\rm s}$ in the derivation leading to our estimate of $\dot{M}$
(eq.~[$\ref{eq:eddr_analytic}$]).  Note, however, that for low mass
SMBHs, $\dot{M}_{\rm Ia}$ generally exceeds the maximum accretion rate
for a thermally stable flow $\dot{M}_{\rm TI}$
(eq.~[\ref{eq:Mdotmax}]), implying that dynamical `blow-out' from SNe
Ia cannot alone prevent cooling instabilities in the nuclei of low
mass galaxies {\bf AG--careful one needs a separate criterion for
  thermal stability when heating is dominated by the Ias}.

The deep gravitational potential wells of high mass galaxies $\sigma_0
\gg f v^{\rm Ia}_{w}$ prevent SN Ia from dynamically clearing out gas
in these systems.  Equation (\ref{eq:eddr_Ia}) nevertheless represents
a cap on the accretion rate in practice because if other source of
heating are ineffective, then the Ia heating rate (eq.~[\ref{eq:vIa}])
is usually sufficiently high to prevent the stagnation radius
(calculated including the heating from SN Ia) from substantially
exceeding $r_{\rm Ia}$.

If Ia supernovae are the dominant source of heating we expect $\rs\sim
r_{\rm Ia}$. If the stagnation radius moves inwards from $\rIa$, the
decreased heating will force it outwards again. If the it moves
outwards the Ia heating will force it inwards. Thus,
equations~\eqref{eq:cooling2} and ~\eqref{eq:Mdotmax} will no longer
be valid as they assume that $\rs$ follows equation
\eqref{eq:stag_simple}. We derive $\dot{q}_{\rm
  heat}/\left|\dot{q}_{\rm rad}\right|$ for $\rs=\rIa$ to evaluate
thermal stability in SNe Ia dominate the heating.

\begin{align}
&\rho(\rIa)\simeq \frac{3 q(\rIa)}{2-\Gamma} \nonumber\\
&T(\rIa)\simeq \frac{\gamma-1}{\gamma}\frac{\mu m_p v^{\rm Ia \,\,
    2}_{w}}{2k}\nonumber\\
&\left.\frac{\dot{q}_{\rm heat}}{\left|\dot{q}_{\rm rad}\right|}\right|_{\rIa}\simeq
\begin{cases}
  90 \Mbheight^{1.3} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{-3}
  \eta_{0.02}^{-2.7} & \text{core}\nonumber\\
   90 \Mbheight^{0.8} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{-2.7}
  \eta_{0.02}^{-2.7} & \text{cusp}\nonumber\\
\end{cases}\\&\simeq
\begin{cases}
  90 \Mbheight^{1.3} (\epsilon_{\rm Ia}/0.4)^{1.7}
  (\tau_{\star}/\th)^{0.5}
 & \text{core}\\
   90 \Mbheight^{0.8} (\epsilon_{\rm Ia}/0.4)^{1.7} (\tau_{\star}/\th)^{0.9}
  & \text{cusp}\\
\end{cases}
\end{align}
